---
title: "Homework 2"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem1

Read Mr Trash Data set, 

```{r}

trashwheel_df =
  read_xlsx ("./data/Trash_data.xlsx",
           sheet = "Mr. Trash Wheel", 
           range= cell_cols("A:N")) %>%
  janitor:: clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sport_balls = round(sports_balls), 
    sports_balls = as.integer(sports_balls)
  )
trashwheel_df
```

Read and clean Precipitation data 2017 2018

```{r}
precip_2018=
   read_xlsx ("./data/Trash_data.xlsx",
              sheet = "2018 Precipitation", 
              skip = 1
              ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year=2018) %>%
  relocate(year)
  precip_2018
  
  
  precip_2017=
   read_xlsx ("./data/Trash_data.xlsx",
              sheet = "2017 Precipitation",
              skip = 1
              ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year=2017) %>%
  relocate(year)
  precip_2017
```

Combine datasets on annual precipitation

```{r}
precip_df= 
  bind_rows(precip_2018, precip_2017) 

month_df = 
  tibble(
    month=1:12, 
    month_name =month.name
  )
month_df

precip_df

left_join(precip_df, month_df, by="month")
```

The dataset information is from the Mr.Trashwheel collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. the dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. 

