---
title: "Homework 2"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem1

Read Mr Trash Data set, 

```{r}

trashwheel_df =
  read_xlsx ("./data/Trash_data.xlsx",
           sheet = "Mr. Trash Wheel", 
           range= cell_cols("A:N")) %>%
  janitor:: clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sport_balls = round(sports_balls), 
    sports_balls = as.integer(sports_balls)
  )
trashwheel_df
```

Read and clean Precipitation data 2017 2018

```{r}
precip_2018=
   read_xlsx ("./data/Trash_data.xlsx",
              sheet = "2018 Precipitation", 
              skip = 1
              ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year=2018) %>%
  relocate(year)
  precip_2018
  
  
  precip_2017=
   read_xlsx ("./data/Trash_data.xlsx",
              sheet = "2017 Precipitation",
              skip = 1
              ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year=2017) %>%
  relocate(year)
  precip_2017
```

Combine datasets on annual precipitation

```{r}
precip_df= 
  bind_rows(precip_2018, precip_2017) 

month_df = 
  tibble(
    month=1:12, 
    month_name =month.name
  )
month_df

precip_df

left_join(precip_df, month_df, by="month")
```

The dataset information is from the Mr.Trashwheel collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. the dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. 

## Problem 2

```{r}
transit_df = 
  read_csv("./data2/NYC_Transit.csv") %>%
  janitor::clean_names() %>%
  select(line,station_name, station_latitude, station_longitude, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, entry, vending, entrance_type, ada )  %>%
mutate(entry = recode(entry, "YES" ="TRUE", "NO" = "FALSE"))
  transit_df
  
```

The dataset contains information from the transit system in NYC. The variables `r variable.names(transit_df)`. In order to clean the data I used the Janitor function to turn the variable names into snakecase I then selected for our variables of interest in order to eliminate unwanted clutter variables. The dataset is `r nrow(transit_df)` rows and `r ncol(transit_df)` columns. The data does not seem tidy given the routes seperate variables and the names are not in snake case. 

There are `r select(distinct(transit_df, line, station_name))%>%
count()` stations and `r filter(transit_df, ada == "TRUE") %>%
distinct(line, station_name)%>%
  count()` are ADA compliant. The proportion of station entrances / exits without vending allow entrance is `r   filter(transit_df, vending == "NO" & entry == "TRUE") %>%
  distinct(line, station_name)%>%
  count() /
     filter(transit_df, vending == "NO") %>%
    distinct(line, station_name)%>%
  count() ` 
  
```{r}
transit_tidy_df = 
  transit_df %>%
  mutate_at(vars(route8:route11), as.character) %>%
  pivot_longer(
      route1:route11, 
      names_to = "route", 
      values_to = "route_name"
      ) %>%
  drop_na(route)

filter(transit_tidy_df,route_name == "A") %>%
  distinct(line, station_name) %>%
 count()

filter(transit_tidy_df, ada == "TRUE" & route_name == "A" ) %>%
  distinct(line, station_name) %>%
  count()

transit_tidy_df

```
  
There are `r filter(transit_tidy_df,route_name == "A") %>%
  distinct(line, station_name) %>% count()` distinct stations that serve the A train. 
`r filter(transit_tidy_df, ada == "TRUE" & route_name == "A" ) %>% distinct(line, station_name) %>% count()` of the stations that serve the A train are ADA compliant. 
 
 
```{r}
pols_df = 
  read_csv("./data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate (mon, into = c("year", "month", "day")) %>%
mutate(month = month.name[as.factor(month)]) %>%
mutate(year = as.numeric(year)) %>%
mutate(president = case_when(prez_dem == 1 ~ "dem", prez_gop ==1 ~ "gop")) %>%
select(-day, -prez_dem, -prez_gop)
```

Arrange function sorts it from lowest to highest and relocate function brings variable to front of df 

```{r}
snp_df = 
  read_csv("./data/snp.csv") %>%
   janitor::clean_names() %>%
  separate (date, into = c("month", "day", "year")) %>%
  mutate(year = as.numeric(year)) %>%
  arrange(year)  %>%
  arrange(month)  %>%
  relocate (year, month)
```

```{r}
unemployment_df = 
  read.csv("./data/unemployment.csv") %>%
  janitor::clean_names() %>%
  mutate(year = as.numeric(year)) %>%
  pivot_longer(
    jan:dec,
     names_to = "month", 
      values_to = "unemployment"
      ) 

```

Merge the datasets 

```{r}
afinal_df = left_join(pols_df, snp_df, by = c("year", "month"))
final_df = left_join(afinal_df, unemployment_df, by = c("year", "month"))
final_df

summarize(final_df, min(year))
summarize(final_df, max(year))
```

The dataset contains information related to the number of national politicians who are democratic or republican at any given time,  unemployment, & Standard & Poor’s stock market index. The variables contained are `r variable.names(final_df)`.The dataset is `r nrow(final_df)` rows and `r ncol(final_df)` columns. The range of years in the data set is `r summarize(final_df, min(year))` and `r summarize(final_df, max(year))`
