---
title: "Homework 2"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem1

Read Mr Trash Data set, 

```{r}

trashwheel_df =
  read_xlsx ("./data/Trash_data.xlsx",
           sheet = "Mr. Trash Wheel", 
           range= cell_cols("A:N")) %>%
  janitor:: clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sport_balls = round(sports_balls), 
    sports_balls = as.integer(sports_balls)
  )
trashwheel_df
```

Read and clean Precipitation data 2017 2018

```{r}
precip_2018=
   read_xlsx ("./data/Trash_data.xlsx",
              sheet = "2018 Precipitation", 
              skip = 1
              ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year=2018) %>%
  relocate(year)
  precip_2018
  
  
  precip_2017=
   read_xlsx ("./data/Trash_data.xlsx",
              sheet = "2017 Precipitation",
              skip = 1
              ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year=2017) %>%
  relocate(year)
  precip_2017
```

Combine datasets on annual precipitation

```{r}
precip_df= 
  bind_rows(precip_2018, precip_2017) 

month_df = 
  tibble(
    month=1:12, 
    month_name =month.name
  )
month_df

precip_df

left_join(precip_df, month_df, by="month")
```

The dataset information is from the Mr.Trashwheel collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. the dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. 

## Problem 2

```{r}
transit_df = 
  read_csv("./data2/NYC_Transit.csv") %>%
  janitor::clean_names() %>%
  select(line,station_name, station_latitude, station_longitude, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, entry, vending, entrance_type, ada )  %>%
mutate(entry = recode(entry, "YES" ="TRUE", "NO" = "FALSE"))
  transit_df
  
```

The dataset contains information from the transit system in NYC. The variables `r variable.names(transit_df)`. In order to clean the data I used the Janitor function to turn the variable names into snakecase I then selected for our variables of interest in order to eliminate unwanted clutter variables. The dataset is `r nrow(transit_df)` rows and `r ncol(transit_df)` columns. 

There are `r select(distinct(transit_df, line, station_name))%>%
count()` stations and `r filter(transit_df, ada == "TRUE") %>%
distinct(line, station_name)%>%
  count()` are ADA compliant. The proportion of station entrances / exits without vending allow entrance is `r   filter(transit_df, vending == "NO" & entry == "TRUE") %>%
  distinct(line, station_name)%>%
  count() /
     filter(transit_df, vending == "NO") %>%
    distinct(line, station_name)%>%
  count() ` 